About

### 2020-03-21 about.Rmd
made graph of David Ortiz WAR over time

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

For top high school players especially, the magnitude of this decision is immense. While certain young players will sign for small contracts, eager to begin their professional career, others will forgo multi-million dollar signing bonuses and attend college. For some, the decision pays off handsomely and the prospect signs for more money as a polished college star. For others, they get injured or fail to perform in college and never play professional baseball.

Money is obviously not the only factor in a high school prospect's decision - the team, prospect's health, and family influences among other things can play a significant role. However, given the amount of money and the opportunity many high school players opt to delay, it begs the question of how the decision plays out for these athletes.

Do high school players who sign out of high school have a better chance of making it to the Major Leagues relative to their peers who attend college? What is the relationship between a player's decision of whether to sign and the longevity and production of their professional baseball career? Of drafted high school players that make it to the big leagues, do players that choose to attend college develop and experience more fruitful careers than players that sign immediately out of high school?

@@

```{r}
download.file("https://www.baseball-reference.com/data/war_daily_bat.txt", dest ="war.csv", mode = "wb")
```

@@

With regards to the rest of my data, I found a dataset in section this week with Kaneesha and emailed Preceptor to see if there are any available class funds to purchase it. The data contains data on all draft picks in the MLB draft dating back to 1965. It contains essentially all of the data I need to join to the WAR data to be able to analyze the questions I posed in the description above. Among these variables are player name, whether they signed, whether they were drafted out of high school or college, their age, if they have made it to the major leagues, the level they currently

Once I access this data, I have a handful of ideas for how I will begin to play with the data. For one, I will compare among players that make it to the major leagues, the average production (WAR) of players who turned down the draft out of high school and those who signed immediately by position. In addition, I will look at the average change in draft position for high school players who turn down the draft. Another idea I will examine is the average speed to the majors in terms of years spent in the minor leagues or age. Finally, I will look at how high up organizationally players in each of these two categories are after different numbers of years in the pros or age. I will also look at career earnings for the two categories by position to see if the decision pays off on average strictly from a financial standpoint.

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidycensus)
library(ggthemes)
library(ggplot2)
library(janitor)
library(dplyr)
library(gt)
library(readxl)
library(sf)
library(scales)
library(magrittr)
library(haven)
library(infer)
library(Lahman)
library(xml2)
library(rvest)
```

@@

  For top high school players especially, the magnitude of this decision is
immense. While certain young players will sign for small contracts, eager to
begin their professional career, others will forgo multi-million dollar signing
bonuses and attend college. For some, the decision pays off handsomely and the
prospect signs for more money as a polished college star. For others, they get
injured or fail to perform in college and never play professional baseball.
Money is obviously not the only factor in a high school prospect's decision -
the team, prospect's health, and family influences among other things can play a
significant role. However, given the amount of money and the opportunity many
high school players opt to delay, it begs the question of how the decision plays
out for these athletes.

Do high school players who sign out of high school have a better chance of
making it to the Major Leagues relative to their peers who attend college? What
is the relationship between a player's decision of whether to sign and the
longevity and production of their professional baseball career? Of drafted high
school players that make it to the big leagues, do players that choose to attend
college develop and experience more fruitful careers than players that sign
immediately out of high school? POSSIBLE RESEARCH proportion of picks on high
school players for each MLB team and what proportion make it to the bigs to see
if certain teams better at developing hs talent

@@

download.file("https://www.baseball-reference.com/data/war_daily_bat.txt", dest ="war.csv", mode = "wb")

```{r, include=FALSE}
csv <- read_csv("war.csv")
ortiz <- csv %>%
  filter(name_common == "David Ortiz")
```

```{r, echo=FALSE}

ggplot(data = ortiz, mapping = aes(age, WAR, color = team_ID)) +
  geom_point() +
  theme_classic() + 
  labs(
    x = "Age",
    y = "Wins Above Replacement",
    title = "David Ortiz Wins Above Replacement by Age",
    subtitle = "Note his jump in WAR upon joining the Red Sox"
  )

```

@@

I have a handful of ideas for how I will begin to play with the data. For one, I will compare among players that make it to the major leagues, the average production (WAR) of players who turned down the draft out of high school and those who signed immediately by position. In addition, I will look at the average change in draft position for high school players who turn down the draft. Another idea I will examine is the average speed to the majors in terms of years spent in the minor leagues or age. Finally, I will look at how high up organizationally players in each of these two categories are after different numbers of years in the pros or age. I will also look at career earnings for the two categories by position to see if the decision pays off on average strictly from a financial standpoint.
